Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-07-24 demystify.tex
updates

The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences). 
Students from top schools with strong research cultures and experiences might already know most of the information given here.
My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
I also hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.

@@

\paragraph{Not a good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.  This is good for you, as you don't want to be in a program where no one can advise you.

@@

In most cases adcom members, even those reviewing the same application, make decision independently and do not talk to each other.
However, in some rare cases we do talk to each other.  For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.  We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser. 
If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.

@@

We only review those that are assigned to us. Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).

For example, I only review those that are assigned to me, which is typically around 50--70 applications for the Fall admission. 
I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam). But usually, even if I look at them, I do not get involved in their evaluation.

@@

@@

\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US). However, \emph{do not} install or use them on university computers, even those given to you by your adviser.  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.  If you need to purchase software,  ask your adviser or the department (\autoref{sec:buying-equipment}).

>>>>

The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences). 
Students from top schools with strong research cultures and experiences might already know most of the information given here.
My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.

@@

\paragraph{Not a good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.  This is good for you, as you don't want to be in a program where no one can advise you.

@@

In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other. It's not a strict rule but a common practice to avoid bias and ensure fairness (and also we don't have time to talk to each other).

However, in some rare cases we do talk to each other.  For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.  We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser. 
If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.

@@

We only review those that are assigned to us, which are already too many. Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).

For example, I only review those that are assigned to me, which is typically around 50--70 applications for the Fall admission. 
I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam). However, even if I look at them, I usually do not get involved in their evaluation.

@@

Keep in mind that these computers would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).  You might also have to return them when you graduate.

@@

\section{Illegal Software}\label{sec:illegal-software}
 Using illegal/cracked software is very common in many countries (and even in the US). However, \emph{do not} install or use them on university computers, even those given to you by your adviser.  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.  If you need to purchase software,  ask your adviser or the department (\autoref{sec:buying-equipment}).
